Reference model in feature selection problems \vspace{.1in}

### 2019-05-23 paper/ref_approach.tex
more corrections on the introduction

Regardless which kind of selection problem we are trying to solve, we consider examples of the reference model approach all those methods that make use of a full-encompassing model, i.e. the reference one. Such an idea is not new to the model reduction literature, it has been used for many years using different type of names. \cite{book:harrell} refers to the reference model as a full model that can be thought as the ``gold standard'' and shows some examples of how it can be used to look for a sparser approximation. He highlights some of the benefits as the possibility to calculate the accuracy with which the submodel approximate the best model and the inheritance of the shrinkage when it is applied to the full model. \cite{paper:faraggi_nn} deal with the necessity of identifying interpretable risk groups in the context of survival data. Neural networks typically perform very well in terms of prediction, but it is not very clear how to interpret the significance of each covariate. \cite{paper:faraggi_nn} propose to combine the benefits of the neural networks, which become what we call reference models, with the interpretability of regression trees, in order to provide a framework to make inferences about risk groups. \cite{paper:paul_preconditioning}, using the term preconditioning, explore approximating models fitting Lasso \citep{paper:tibshirani_lasso} or stepwise regression against consistent estimates $\hat{y}$ of a reference model instead of the observed response quantities $y$. \cite{book:harrell} includes also the possibility of using simple least squares for the approximating models as long as the target is the linear predictor of a full regression model \hl{CHECK}.
\\
The reference model approach has been used for long time also in the Bayesian framework, an example the pioneering work by \cite{paper:reference_lindley}. In the case of generalised linear models (GLMs), \cite{paper:tran_predictivelasso} provide a procedure to estimate and do variable selection with the goal of prediction. The best approximating model is recovered by optimising the Kullback-Liebler divergence between the predictive distribution of the reference model and the approximating model plus an $l_{1}$ penalty on the submodel coefficients. The resulting problem is convex and thus many efficient optimisation algorithm are available. Following \cite{paper:goutis_projection}, \cite{paper:original_proj}  instead project the posterior distribution of the reference model in each candidate model and then select the smallest one with enough explanatory power avoiding any prior elicitation for the submodels. In the special case of candidate models in the family of the GLMs, the optimisation problem correspond to the maximum likelihood estimation problem having the expected model point prediction by the reference model instead of the observed target values. Different extensions have followed \cite{paper:original_proj}, as \cite{paper:nott_projection} that add an $l_{1}$ penalty to select the best approximating model. \cite{paper:projpred} extend the projection algorithm with the clustered projection and choose the best model comparing the predictive performance with the reference model one.

\cite{paper:projpred} show in their experiments the benefits of the projection predictive approach in terms of predictive ability of the selected model and smaller size compared to other methods. It could be of interest to know how often a certain method includes non-relevant features. One could argue that such question can not have a general answer, because it depends both on the data and on the choice of the reference model. Moreover in real settings it is hard to define what relevant means, in most of the cases all the features have some level of relevance in term of prediction ability and thus the question looses of meaning.  However, it is still possible to simulate certain artificial scenarios and control which features are included. \cite{paper:model_selection} report results about the proportion of relevant and non-relevant features selected with different variable selection methods, showing overall better results in favour of the projection predictive approach. The explored submodels using the projection show also a higher performance and stability with respect to the compared methods, given the same submodel size.

In this paper we try to give further examples and intuitions on why the reference model helps. Using a real world dataset, we show the improvements in the selection when using the projection predictive approach versus the stepwise regression. Wherever it has been used, the reference model can be always seen acting as a filter on the observed data. Following this idea, we devise a simple reference model approach that can be applied on top of any selection procedure. Our goal is to show the reference model benefits, regardless of what specific procedure is applied. We take as reference problem the normal means estimation problem. We apply different state-of-the-art methods comparing in each case the results using the data or using a reference model. In case of full Bayesian methods, we report results in terms of shrinkage and accuracy of the estimates, whereas when a full selection is carried out, we compute quantities as false discovery rate, sensitivity and a measure of the stability of the selection. We believe that these results should motivate the use of a reference model when doing model reduction, regardless of the specific method used and thus we hope this work can be of interest to a large number of people.

The paper is structured as follows. In Section \ref{reference-model-approach} we introduce the idea of the reference model, its benefits with examples and how we used it as a filter on our data in a very simple way. In Section \ref{comparison} we use different methods for feature selection comparing the selection results using or not the reference model in the framework of the normal means problem. Finally, in Section \ref{conclusions} we present our conclusions.

>>>>

Regardless which kind of selection problem we are trying to solve, we consider examples of the reference model approach all those methods that make use of a full-encompassing model, i.e. the reference one. Such an idea is not new to the model reduction literature, it has been used for long time using different names. \cite{book:harrell} refers to the reference model as a full model that can be thought as the ``gold standard'' and shows some examples of how it can be used to look for a sparser approximation. He highlights some of the benefits as the possibility to calculate the accuracy with which the submodel approximates the best model and the inheritance of the shrinkage when it is applied to the full model. \cite{paper:faraggi_nn} deal with the necessity of identifying interpretable risk groups in the context of survival data. Neural networks typically perform very well in terms of prediction, but it is not very clear how to interpret the significance of each covariate. \cite{paper:faraggi_nn} propose to combine the benefits of the neural networks, which become what we call reference models, with the interpretability of regression trees, in order to provide a framework to make inferences about risk groups. \cite{paper:paul_preconditioning}, using the term preconditioning, explore approximating models fitting Lasso \citep{paper:tibshirani_lasso} or stepwise regression against consistent estimates $\hat{y}$ of a reference model instead of the observed response quantities $y$. \cite{book:harrell} includes also the possibility of using simple least squares for the approximating models as long as the target is the linear predictor of a full regression model \hl{CHECK}.
\\
The reference model approach has been used for long time also in the Bayesian framework, an example the pioneering work by \cite{paper:reference_lindley}. In the case of generalised linear models (GLMs), \cite{paper:tran_predictivelasso} provide a procedure to estimate and do variable selection with the goal of prediction. The best approximating model is recovered by optimising the Kullback-Liebler divergence between the predictive distribution of the reference model and the approximating model plus an $l_{1}$ penalty on the submodel coefficients. The resulting problem is convex and thus many efficient optimisation algorithm are available. Following \cite{paper:goutis_projection}, \cite{paper:original_proj}  instead project the posterior distribution of the reference model in each candidate model and then select the smallest one with enough explanatory power avoiding any prior elicitation for the submodels. In the special case of candidate models in the family of the GLMs, the optimisation problem correspond to the maximum likelihood estimation problem having the expected model point prediction by the reference model instead of the observed target values. Different extensions have followed \cite{paper:original_proj}, as \cite{paper:nott_projection} that add an $l_{1}$ penalty to select the best approximating model. \cite{paper:projpred} extend the projection algorithm with the clustered projection and choose the best model comparing the predictive performance with the reference model one.

\cite{paper:projpred} show in their experiments the benefits of the projection predictive approach in terms of predictive ability of the selected model and smaller size compared to other methods. It could be of interest to know how often a certain method includes non-relevant features. One could argue that such question can not have a general answer, because it depends both on the data and, when it is used, on the choice of the reference model. Moreover in real settings it is hard to define what relevant means, in most of the cases all the features have some level of relevance in term of prediction ability and thus the question looses of meaning.  However, it is still possible to simulate certain artificial scenarios and control which features are included. \cite{paper:model_selection} report results about the proportion of relevant and non-relevant features selected with different variable selection methods, showing overall better results in favour of the projection predictive approach. The explored submodels using the projection show also a higher performance and stability with respect to the compared methods, given the same submodel size.

In this paper we try to give further examples and intuitions on why the reference model helps. Using a real world dataset, we show the improvements in the selection when using the projection predictive approach versus the stepwise regression. However it has been used, the reference model can be always seen acting as a filter on the observed data. Following this idea, we devise a simple reference model approach that can be applied on top of any selection procedure. Our goal is to show the reference model benefits, regardless of what specific procedure is applied. We take as reference problem the normal means estimation problem. We apply different state-of-the-art methods comparing in each case the results using the data or using a reference model. In case of full Bayesian methods, we report results in terms of shrinkage and accuracy of the posterior estimates, whereas when a full selection is carried out, we compute quantities as false discovery rate, sensitivity and a measure of the stability of the selection. We believe that these results should motivate the use of a reference model when doing model reduction, regardless of the specific method used and, thus, we hope this work can be of interest to a large number of practitioners.

The paper is structured as follows. In Section \ref{reference-model-approach} we introduce the idea of the reference model, its benefits with examples, including the projection predictive approach, and how it can be used as a filter on data in a simple way. In Section \ref{comparison} we use different methods for feature selection comparing the selection results using or not the reference model in the framework of the normal means problem. Finally, in Section \ref{conclusions} we present our conclusions.
